A weighted sampling algorithm to designing RNA sequences with target secondary structure and nucleotide distribution

### 2013-01-08 ISMB/introduction_ISMB.tex
completed title/abstract/intro
typos

Most of RNA functions are, at least partially, encoded by the three-dimensional molecular structures, which are themselves primarily determined by the secondary structures. The development of efficient algorithms for designing RNA sequences with pre-defined secondary structures is thus a milestone to enter the synthetic biology era. \RNAinverse pioneered RNA secondary structure design algorithms. It has been developed and distributed with the Vienna RNA package \cite{Hofacker:1994}. However, only posterior experimental studies revealed the potential and practical impact of these techniques. Thereby, during the last 6 years many improvements and variants of \RNAinverse have been proposed. Conceptually, almost all existing programs follow the same approach. First a seed sequence is selected, then a local search strategy is used to mutate the seed and find in its vicinity a sequence with desired folding properties. Using this strategy, \INFORNA \cite{Busch:2006uq}, \RNASSD \cite{Aguirre-Hernandez:2007kx} and \NUPACK \cite{Zadeh:2011fk} significantly improved the performance of the RNA secondary structure design algorithms. More recent research studies aimed to include more constraints in the selection criteria.\RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \cite{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \cite{Lyngso:2012vn}.

We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \cite{Levin:2012kx}. Instead of a local search approach, we propose to apply a global sampling strategy of the mutational landscape using our \RNAmutants algorithm \cite{Waldispuhl2008}. This methodology offers promising performance but  suffers from a prohibitive running time complexity. Following our work, Garcia-Martin and co-workers proposed with \RNAiFOLD an alternate methodology using constraint programming techniques to prune the mutational landscape. The latter also suffers from prohibitive running times, but it is worth noting that it proposes a seed-less approach to the RNA secondary structure design problem.

In this paper we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}. \ourprog addresses previous limitations of \RNAensign and offers new functionalities. First, while our previous program had a running time complexity of $\mathcal{O}(n^4)$, \ourprog has now a linear-time and space complexity that allows him to demonstrate similar speeds to any local search algorithm. Next, \ourprog is \textit{seed-less}. Unlike \RNAensign, it does not require a seed sequence to start its search. Finally, \ourprog implements a novel algorithm using weighted sampling techniques \cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \gc~content of the solution. This functionality is essential because wild sequences within living organisms often present medium or low \gc~content. Presumably to offer better transcription rates and/or structural plasticity. Previous programs do not allow to control this parameter and tend to output sequences with high \gc~contents.

We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \textsf{RNA STRAND} database \cite{andronescu2008rna}. To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse. Remarkably, our glocal methodology overcome both local and global approaches.

>>>>

Most of RNA functions are, at least partially, encoded by the three-dimensional molecular structures, which are themselves primarily determined by the secondary structures. The development of efficient algorithms for designing RNA sequences with pre-defined secondary structures is thus a milestone to enter the synthetic biology era. \RNAinverse pioneered RNA secondary structure design algorithms. It has been developed and distributed with the Vienna RNA package \cite{Hofacker:1994}. However, only posterior experimental studies revealed the potential and practical impact of these techniques. Thereby, during the last 6 years many improvements and variants of \RNAinverse have been proposed. Conceptually, almost all of existing algorithms follow the same approach. First a seed sequence is selected, then a local search strategy is used to mutate the seed and find, in its vicinity, a sequence with desired folding properties. Using this strategy, \INFORNA \cite{Busch:2006uq}, \RNASSD \cite{Aguirre-Hernandez:2007kx} and \NUPACK \cite{Zadeh:2011fk} significantly improved the performance of RNA secondary structure design algorithms. More recent research studies aimed to include more constraints in the selection criteria.\RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \cite{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \cite{Lyngso:2012vn}.

We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \cite{Levin:2012kx}. Instead of a local search approach, we proposed a global sampling strategy of the mutational landscape based on the \RNAmutants algorithm~\cite{Waldispuhl2008}. This methodology offered promising performances, but suffered from  prohibitive runtime and memory consumption. Following our work, Garcia-Martin and co-workers proposed \RNAiFOLD, an alternate methodology that uses constraint programming techniques to prune the mutational landscape. While also suffering from prohibitive running times, it is worth noting that this latter algorithm also proposes a seed-less approach to the RNA secondary structure design problem.

In this paper, we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}. \ourprog addresses previous limitations of \RNAensign and offers new functionalities. First, while our previous program had a running time complexity of $\mathcal{O}(n^5)$, \ourprog now runs in linear-time and space complexity, allowing it to demonstrate similar speeds as any local search algorithm. Next, \ourprog is \textit{seed-less}. Unlike \RNAensign, it does not require a seed sequence to initiate its search. Finally, \ourprog implements a novel algorithm using weighted sampling techniques~\cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \GCContent of the solution. This functionality is essential because wild sequences within living organisms often present medium or low \GCContent, presumably to offer better transcription rates and/or structural plasticity. Previous programs do not allow to control this parameter and tend to output sequences with high \GCContent{}s.

We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \RNASTRAND database \cite{andronescu2008rna}. To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse. Remarkably, our glocal methodology overcomes both local and global approaches.

### 2013-01-08 ISMB/main_ISMB.tex
typo & supp data
typos

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\farna}{\texttt{FARNA}\xspace}
\newcommand{\mcfoldmcsym}{\texttt{MC-Pipeline}\xspace}
\newcommand{\mcfold}{\texttt{MC-Fold}\xspace}
\newcommand{\mcsym}{\texttt{MC-Sym}\xspace}
\newcommand{\nast}{\texttt{NAST}\xspace}
\newcommand{\ifoldrna}{\texttt{iFoldRNA}\xspace}
\newcommand{\rnafold}{\texttt{RNAfold}\xspace}
\newcommand{\rnasubopt}{\texttt{RNAsubopt}\xspace}
\newcommand{\rnawolf}{\texttt{RNAwolf}\xspace}
\newcommand{\rnastructure}{\texttt{RNAstructure}\xspace}
\newcommand{\contrafold}{\texttt{contrafold}\xspace}
\newcommand{\unafold}{\texttt{unafold}\xspace}
\newcommand{\rnadd}{\texttt{RNA2D3D}\xspace}
\newcommand{\assemble}{\texttt{assemble}\xspace}
\newcommand{\fred}{\texttt{FR3D}\xspace}
\newcommand{\rnajunction}{\texttt{RNAjunction}\xspace}
\newcommand{\rnamotif}{\texttt{RNAmotif}\xspace}
\newcommand{\treefolder}{\texttt{TreeFolder}\xspace}
\newcommand{\barnacle}{\texttt{BARNACLE}\xspace}
\newcommand{\contextfold}{\texttt{contextfold}\xspace}
%\newcommand{\citep}{\cite}
\newcommand{\Z}[2]{\mathcal{Z}{\substack{(#1)\\ [#2]}}}
\newcommand{\Y}[3]{\mathcal{Y}_{\substack{(#1)\\ [#3]}}^{#2}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\Kron}{\delta}
\newcommand{\ub}{\bullet}

@@

\title{A glocal approach to designing RNA sequences with target secondary structure and nucleotide distribution}
\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, ƒcole Polytechnique, Palaiseau, France.\\ \small $^*$Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}

>>>>

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\farna}{\texttt{FARNA}\xspace}
\newcommand{\mcfoldmcsym}{\texttt{MC-Pipeline}\xspace}
\newcommand{\mcfold}{\texttt{MC-Fold}\xspace}
\newcommand{\mcsym}{\texttt{MC-Sym}\xspace}
\newcommand{\nast}{\texttt{NAST}\xspace}
\newcommand{\ifoldrna}{\texttt{iFoldRNA}\xspace}
\newcommand{\rnafold}{\texttt{RNAfold}\xspace}
\newcommand{\rnasubopt}{\texttt{RNAsubopt}\xspace}
\newcommand{\rnawolf}{\texttt{RNAwolf}\xspace}
\newcommand{\rnastructure}{\texttt{RNAstructure}\xspace}
\newcommand{\contrafold}{\texttt{contrafold}\xspace}
\newcommand{\unafold}{\texttt{unafold}\xspace}
\newcommand{\rnadd}{\texttt{RNA2D3D}\xspace}
\newcommand{\assemble}{\texttt{assemble}\xspace}
\newcommand{\fred}{\texttt{FR3D}\xspace}
\newcommand{\rnajunction}{\texttt{RNAjunction}\xspace}
\newcommand{\rnamotif}{\texttt{RNAmotif}\xspace}
\newcommand{\treefolder}{\texttt{TreeFolder}\xspace}
\newcommand{\barnacle}{\texttt{BARNACLE}\xspace}
\newcommand{\contextfold}{\texttt{contextfold}\xspace}

\newcommand{\RNASTRAND}{{\sf RNA STRAND}\xspace}

%\newcommand{\citep}{\cite}
\newcommand{\Z}[2]{\mathcal{Z}{\substack{(#1)\\ [#2]}}}
\newcommand{\Y}[3]{\mathcal{Y}_{\substack{(#1)\\ [#3]}}^{#2}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\Kron}{\delta}
\newcommand{\ub}{\bullet}

@@

\title{A weighted sampling algorithm to designing RNA sequences with target secondary structure and nucleotide distribution}
\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.\\ \small $^*$Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}

### 2013-01-08 ISMB/methods_ISMB.tex
typos

We introduce a probabilistic model for the design of RNA sequences with a specific $\Gb+\Cb$ content and folding into a predefined secondary structure.
For the sake of simplicity, we choose to base this proof-of-concept implementation on a simplified free-energy function $\ES(\cdot)$, which only considers the contributions of 
stacked canonical base-pairs.

We show how a modification of the dynamic programming scheme introduced in the previously contributed\RNAmutants allows for the sampling of good and diverse design candidates, in linear time and space complexities.

@@

\subsubsection{Simplified energy model}
The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}). Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
\begin{align*}
  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
\end{align*}
where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e. not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$). This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = \infty$).

@@

Consider the {\bf covering base-pair} $\CNBP{i}$ of a position $i$, defined as the base-pair $(i',j')$ such that $i'<i<j'$ and which minimizes $i-i'$. For the sake of simplicity, and without loss of generality, let us assume that there exists a virtual base-pair $(-1,n)$ in $\Target$, so that any position in $[0,n-1]$ has a covering base-pair.
Then one easily shows by induction that, starting from the complete interval $[0,n-1]$, the recursive calls only evaluate $\mathcal{Z}$ over intervals $[i,j]$ such that $j=j'-1$, $\CNBP{i}=[i',j']$.

\CFramed{This property clearly holds initially for the complete interval $[0,n-1]$, owing to the virtual base-pair $(-1,n)$. Furthermore, if it holds for a given $(i,j)$, $\CNBP{i}=[i',j+1]$, then:
\begin{itemize} 
  \item If $i$ is unpaired (Case 1), then $\CNBP{i+1}=\CNBP{i}=[i',j+1]$, and the property is satisfied by the subsequent call over $[i+1,j]$. 
  \item If $i$ is paired with $j$ and stacking onto $(i-1,j+1)$ (Case 2), then one has $\CNBP{i+1}=[i,j]$ and the  subsequent call over $[i+1,j-1]$ satisfies the property.
  \item If $i$ is paired with some $k$ (Case 3), then $\CNBP{i+1}=[i,k]$ and $\CNBP{k+1}=[i',j+1]$. Therefore the subsequent calls, respectively over $[i+1,k-1]$ and $[k+1,j]$, both satisfy the property.
\end{itemize}
The case of an empty sequence does not cause a recursive call and must not be verified.
}
Since, in any interval $[i,j]$ satisfying the property, the ending position $j$ is completely determined by $i$, then there exists at most $\Theta(|\Target|)$ intervals that need to be evaluated, and we conclude on the overall complexity with the remark that $a$, $b$, $a'$ and $b'$ vary over bounded ranges and only contribute to a constant factor in the complexity.
\end{proof}

@@

Since, during each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position, and that the number of executions of the loops are bounded by a constant, then the complexity of the algorithm grows linearly on the targeted length.

@@

Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent. Such a value can be determined using any numerical recipe, and a simple bisection procedure, further described in earlier work~\cite{Waldispuhl2011}, is used in our implementation. Furthermore, the concentration of the distribution, as asserted by the limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp. $\Theta(1)$, i.e. constant) for the expected number of attempts before generating a sequence having the exact targeted \GCContent  (resp. within $[gc-\kappa,gc+\kappa]$, where $\kappa\in\Theta(1/\sqrt n)$).

\subsection{Redesigning unpaired regions: A local/global hybrid approach}
Due to the simplification in our energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest similarities of refolded design candidates with the targeted structure. To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions, specifying a constraint mask to prevent stacking base-pairs from being modified.
{\em todo: discuss evolution of \GCContent}

>>>>

We introduce a probabilistic model for the design of RNA sequences with a specific $\Gb+\Cb$ content and folding into a predefined secondary structure.
For the sake of simplicity, we choose to base this proof-of-concept implementation on a simplified free-energy function $\ES(\cdot)$, which only considers the contributions of 
stacked canonical base-pairs. We show how a modification of the dynamic programming scheme used in the \RNAmutants allows for the sampling of good and diverse design candidates, in linear time and space complexities.

@@

\subsubsection{Simplified energy model}
The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}). Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
\begin{align*}
  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
\end{align*}
where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble pairs (i.e. not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$). This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = \infty$).

@@

Consider the {\bf covering base-pair} $\CNBP{i}$ of a position $i$, defined as the base-pair $(i',j')$ such that $i'<i<j'$ and which minimizes $i-i'$. For the sake of simplicity, and without loss of generality, let us assume that there exists a virtual base-pair $(-1,n)$ in $\Target$, so that any position in $[0,n-1]$ has a covering base-pair.
Then, it is easily shown by induction that, starting from the complete interval $[0,n-1]$, the recursive calls only evaluate $\mathcal{Z}$ over intervals $[i,j]$ such that $j=j'-1$, $\CNBP{i}=[i',j']$. 
\CFramed{This property clearly holds initially for the complete interval $[0,n-1]$, owing to the virtual base-pair $(-1,n)$. Furthermore, if it holds for a given $(i,j)$, $\CNBP{i}=[i',j+1]$, then:
\begin{itemize} 
  \item If $i$ is unpaired (Case 1), then $\CNBP{i+1}=\CNBP{i}=[i',j+1]$, and the property is satisfied by the subsequent call over $[i+1,j]$. 
  \item If $i$ is paired with $j$ and stacking onto $(i-1,j+1)$ (Case 2), then one has $\CNBP{i+1}=[i,j]$ and the  subsequent call over $[i+1,j-1]$ satisfies the property.
  \item If $i$ is paired with some $k$ (Case 3), then $\CNBP{i+1}=[i,k]$ and $\CNBP{k+1}=[i',j+1]$. Therefore the subsequent calls, respectively over $[i+1,k-1]$ and $[k+1,j]$, both satisfy the property.
\end{itemize}
The case of an empty sequence does not cause a recursive call and must not be verified.
}
Since, in any interval $[i,j]$ satisfying the property, the ending position $j$ is completely determined by $i$, then there exists at most $\Theta(|\Target|)$ intervals that need to be evaluated, and we conclude on the overall complexity with the remark that $a$, $b$, $a'$ and $b'$ vary over bounded ranges and only contribute to a constant factor in the complexity.
\end{proof}

@@

Since, on each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position, and that the number of executions of the loops are bounded by a constant, then the complexity of the algorithm grows linearly on the targeted length.

@@

Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent. Such a value can be determined using any numerical recipe, and a simple bisection procedure, further described in earlier work~\cite{Waldispuhl2011}, is used in our implementation. Furthermore, the concentration of the distribution, asserted by its limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp. $\Theta(1)$, i.e. constant) for the expected number of attempts before generating a sequence having the exact targeted \GCContent  (resp. within $[gc-\kappa,gc+\kappa]$, where $\kappa\in\Theta(1/\sqrt n)$).

\subsection{Postprocessing unpaired regions: A local/global hybrid approach}
\label{subsec:glocal_method}
Due to our simplified our energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest similarities of refolded design candidates with the targeted structure. To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions, specifying a constraint mask to prevent stacking base-pairs from being modified and, if necessary, reetablishing stacking-pair content afterward.
{\em todo: discuss evolution of \GCContent}

### 2013-01-08 ISMB/results_ISMB.tex
add section RNAinverse
typos

\subsection{Implementation}
\label{sec:implementation}
Our software, \texttt{Incarnation}, was implemented in Python2.7. We used
\emph{RNAinverse} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
All time benchmarks were run on a one AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.
 We used as parameter $\beta=15$, the penalty for an invalid base pair.

We present the average times spent by \texttt{Incarnation} and \RNAinverse to generate one sequence
with the required \texttt{C+G} content in Fig.~\ref{fig:time}. As expected, the time grows linearly
in function of the length of the structures for \texttt{Incarnation}.  A highly interesting feature
is the huge decrease in time of computation when there is $15\%$ free 
nucleotides inside the loops of the secondary structure.

@@

\subsection{Dataset}
To evaluate the quality of our method, we used secondary
structures from the \textit{RNA STRAND} database~\cite{andronescu2008rna}.
Those are known secondary structure from a variety of organisms.
We select the same subset of $50$ structures as in~\cite{Levin:2012kx}, 
which have lengths between $20$ and $100$ nucleotides.

 To ease the visualization of the data, we clustered together the strucures
 in function of their length, stacks density and proportion of free 
 nucleotides in loops. We present in Fig.~\ref{fig:bins} the distribution
 of the structures in the different bins.

@@

\subsection{Design}
 To benchmark our method, we first started by sampling a minimum
 $100$ sequence per structures. After correcting the free nucleotides with
 \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.

A first criterion is which percentage of sequences have as MFE the target
secondary structures. We are also interested by the number of structures
for which we can design at least one sequence with the desired MFE.
We show the results when using only \texttt{Incarnation} in 
 Fig.~\ref{fig:mfe_struct_solved_noinverse}. The high rate of failure
is due to the fact that no selection criterions are applied to
free nucleotides. A local strategy is thus needed.
After processing \texttt{Incarnation} sequences with \RNAinverse to 
locally optimize the free nucleotides, we obtain the results 
in Fig.~\ref{fig:mfe_struct_solved}. We observe
that length is not a good predictor for the quality of the results. Instead,
a high number of free nucleotides in the structure seems to be a 
good measure of its design hardness. 
On top of that, with a \texttt{C+G} content of $50\%$ and more, at least
one solutions was found for almost all structures. The difficulty of 
designing sequences for targets with a high number of free nucleotides 
 in loops appears clearly in the last subfigure. Even with a high \texttt{C+G} content a our method was not able to generate solutions for a few
 structures.

@@

To evaluate the global quality of \texttt{Incarnation} sequences, we show
in Fig.~\ref{fig:ss_sens} the ratio of well predicted base pairs in the
MFE structure of our sampled sequences. We can observe that, in all cases,
the hardest sequence to design are those with an extremely low \texttt{C+G}
content. As anticipated, those are the sequences with the weakest bonds.
We notice that the most accurate sequences yield a \texttt{C+G} content
of $70\pm 10\%$.

@@

In this section, we are only interested  in the samples folding into the 
targetMFE.

A desirable feature in sequence design, is to produce samples with a high
diversity. We show in Fig.~\ref{fig:nb_sols_entropy} the number of correct
solutions as their average entropy and base pair entropy, since 
\texttt{Incarnation}  only influences the distribution of nucleotides inside 
stacked base pairs. The most constraining factor for generating valid
 solutions seems to be  low \texttt{C+G} content and the percentage of free nucleotides. Our method is able to generate

Also, a critical properties in RNA sequence design is 
the frequency of the MFE. 
The sequence should be in its target conformation long enough to
perform the desired action. Presented in Fig.~\ref{fig:freq} we see that
there is a slow decline of the frequency with the increase in size. Yet,
for an average \texttt{C+G} content, the frequency stays over $10\%$ even
at size of $100$ nucleotides.

@@

>>>>

\subsection{Implementation}
\label{sec:implementation}
Our software, \texttt{Incarnation}, was implemented in {\tt Python~2.7}. We used
\RNAinverse from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
All time benchmarks were run on a one AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.
 We used as parameter $\beta=15$, the penalty for an invalid base pair.

We present the average times spent by \texttt{Incarnation} and \RNAinverse to generate one sequence
with the required \texttt{C+G} content in Figure~\ref{fig:time}. As expected, the time grows linearly
in function of the length of the structures for \texttt{Incarnation}.  A highly interesting feature
is the huge decrease in time of computation when there is $15\%$ free 
nucleotides inside the loops of the secondary structure.

@@

\subsection{Dataset}
To evaluate the quality of our method, we used secondary
structures from the \RNASTRAND database~\cite{andronescu2008rna}.
Those are known secondary structure from a variety of organisms.
We select the same subset of $50$ structures as in~\cite{Levin:2012kx}, 
which have lengths between $20$ and $100$ nucleotides.

 To ease the visualization of the data, we clustered together the strucures
 in function of their length, stacks density and proportion of free 
 nucleotides in loops. We present in Figure~\ref{fig:bins} the distribution
 of the structures in the different bins.

@@

\subsection{Design}
 To benchmark our method, we first started by sampling a minimum
 of $100$ sequence per structure. After correcting the free nucleotides with
 \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.

A first criterion is which percentage of sequences have as MFE the target
secondary structures. We are also interested by the number of structures
for which we can design at least one sequence with the desired MFE.
We show the results when using only \texttt{Incarnation} in 
 Figure~\ref{fig:mfe_struct_solved_noinverse}. As mentioned earlier, the high rate of failure
is due to the fact that no selection criterion is applied to
unpaired nucleotides. A local strategy is thus needed.
After processing \texttt{Incarnation} sequences with \RNAinverse to 
locally optimize the free nucleotides, we obtained the results 
shown in Figure~\ref{fig:mfe_struct_solved}. We observed
that length is not a good predictor for the quality of the results. Instead,
a high number of free nucleotides in the structure seems to be a 
good measure of its design hardness. 
On top of that, with a \texttt{C+G} content of $50\%$ and more, at least
one solutions was found for almost all structures. The difficulty of 
designing sequences for targets with a high number of free nucleotides 
 in loops appears clearly in the last subfigure. Even with a high \texttt{C+G} content a our method was not able to generate solutions for a few
 structures.

@@

To evaluate the global quality of \texttt{Incarnation} sequences, we show
in Figure~\ref{fig:ss_sens} the ratio of well predicted base pairs in the
MFE structure of our sampled sequences. We can observe that, in all cases,
the hardest sequence to design are those with an extremely low \texttt{C+G}
content. As anticipated, those are the sequences with the weakest bonds.
We notice that the most accurate sequences yield a \texttt{C+G} content
of $70\pm 10\%$.

@@

In this section, we further analyze the generated sequences that fold into the 
target structure.

A desirable feature in sequence design, is to produce samples with a high
diversity. We show in Figure~\ref{fig:nb_sols_entropy} the number of correct
solutions as their average entropy and base pair entropy, since 
\texttt{Incarnation}  only influences the distribution of nucleotides inside 
stacked base pairs. The most constraining factor for generating valid
 solutions seems to be  low \texttt{C+G} content and the percentage of free nucleotides. Our method is able to generate

Also, a critical properties in RNA sequence design is 
the frequency of the MFE. 
The sequence should be in its target conformation long enough to
perform the desired action. Presented in Figure~\ref{fig:freq} we see that
there is a slow decline of the frequency with the increase in size. Yet,
for an average \texttt{C+G} content, the frequency stays over $10\%$ even
at size of $100$ nucleotides.

@@

\subsection{Global sampling, Local search and Glocal approach}

\label{fig:rnainverse}
